---
title: 'Assignment 5 STAT 315-463:  Multivariable Statistical Methods and Applications'
author: 'Lisa Lu 31088272'
header-includes:
   - \usepackage{bm}
   - \usepackage{amsmath}
output: 
  pdf_document:
    number_sections: no
---


```{r}
# Read in data file
initobs <- read.table("initobs.csv", header = TRUE, sep = ',', na.strings = "na")
```

**a):Explain why observer should be included in a model as a random effect.**

The observers are the source of random variation. It is assumed here that these 
observers are from a random sample selected from a large collection of observers.

**b) What distribution would be appropriate for these data and why?**
```{r}
plot(as.factor(initobs$Trt), initobs$Bc/initobs$Steps, xlab="Treatment",ylab="Prob of having weed")
```
These data are not continuous, therefore it is not very suitable to use normal linear regression models. 


**c) Fit an appropriate random effects model to these data**

* a. Discuss the results of the analysis, include comments about the following:
  + i). Scaled residuals
  + ii). Random effects
  + iii). Fixed effects
* b. What would your overall conclusion be?

**d) Draw graphs of the residuals and the random effects and comment on these.**
