---
title: 'Lab 4 STAT 315-463:  Multivariable Statistical Methods and Applications'
header-includes:
   - \usepackage{bm}
   - \usepackage{amsmath}
output: 
  pdf_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Thursday 16 March 2023**  
\vspace{5 mm}

## The 'Data'

For this lab, the data is a black and white photo of Karl Perason, which is available from Learn. You will the code below helpful 

```{r ,echo=TRUE, eval=FALSE}
library(jpeg)
# Importing Karl Pearson
pic <- readJPEG("Pearsonphoto.jpg")

#As the imported image appears as an array, the data we will use will be the first part of the array 
picdata<-pic[,,1]
```


## What to do.

Explore various methods of dimension reduction and clustering as looked at in the lectures in Weeks 3 and  4. These are:

- Principal Component Analysis (PCA). 

  - When looking at PCA, you may find it interesting to see how well a low dimensional approximation of the data can replicate the photo. If you wish to re-export the image, use the function \texttt{writeJPEG(lddata, target="ldimage.jpeg", quality=1)}, where \texttt{lddata} is the matrix containing the low dimensional representation of the data, and \texttt{ldimage.jpeg} is the name of the exported image.

- hierarchical clustering 
- K-means clustering
 - Clustering based on multivariate normal mixtures.  

    - For the three clustering techniques, it might be interesting to investigate if the clusters determined correspond to meaningful features in the photo.

