---
title: "Assignment 6 STAT 315-463: Multivariable Statistical Methods and Applications"
author: "Lisa Lu 31088272"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# QUESTION 1 Generalised additive models

a) Import the data to R and fit a series of GAMs to the Value using a smoother on Date.
```{r}
# Read in the datasets and convert the string "Date" into Date datatype variables
CCC05 <- read.table("CCC05.csv", header = TRUE, sep = ',', na.strings = "na")
CCC05$Date <- as.Date(CCC05$Date, "%d/%m/%Y")
```


```{r, message=FALSE, warning=FALSE}
library(gam)
library(ggplot2)

# Kept showing Error in names(dat) <- object$term : 
# 'names' attribute [1] must be the same length as the vector [0]
CCC05 <- transform(CCC05, ndate = as.numeric(Date),
                   nyear  = as.numeric(format(Date, '%Y')),
                   nmonth = as.numeric(format(Date, '%m')),
                   nday = as.numeric(format(Date, '%j')))

# Start with the default model and 4 more with different spar parameters

CCC05.gam <- gam(Value ~ s(nyear) + s(nmonth) +s(nday), data = CCC05 )
pred_default <- predict(CCC05.gam)
summary(CCC05.gam)

CCC05.gam1 <- gam(Value ~ s(nyear, sp=1.0) + s(nmonth, sp = 1.0) +
                    s(nday, sp = 1.0), data = CCC05)
pred1 <- predict(CCC05.gam1)
summary(CCC05.gam1)

CCC05.gam2 <- gam(Value ~ s(nyear, sp=0.01) + s(nmonth, sp = 0.01) +
                    s(nday, sp = 0.01), data = CCC05)
pred2 <- predict(CCC05.gam2)
summary(CCC05.gam2)

CCC05.gam3 <- gam(Value ~ s(nyear, sp=0.25) + s(nmonth, sp = 0.25) +
                    s(nday, sp = 0.25), data = CCC05)
pred3 <- predict(CCC05.gam3)
summary(CCC05.gam3)

CCC05.gam4 <- gam(Value ~ s(nyear, sp=0.75) + s(nmonth, sp = 0.75) +
                    s(nday, sp = 0.75), data = CCC05)
pred4 <- predict(CCC05.gam4)
summary(CCC05.gam4)

ggplot(data = CCC05, aes(x = Date, y = Value)) +
  geom_point() +
  geom_line(aes(x=Date, y=pred_default), colour = "red") +
  geom_line(aes(x=Date, y=pred1), colour = "orange") +
  geom_line(aes(x=Date, y=pred2), colour = "yellow") +
  geom_line(aes(x=Date, y=pred3), colour = "green") +
  geom_line(aes(x=Date, y=pred4), colour = "blue")
```
```{r}
data <- data.frame(spar0 = fitted.values(CCC05.gam),
                   spar1.0 = fitted.values(CCC05.gam1),
                   spar0.01 = fitted.values(CCC05.gam2),
                   spar0.25 = fitted.values(CCC05.gam3),
                   spar0.75 = fitted.values(CCC05.gam4)
                   )

boxplot(data)

par(mfrow=c(2,2))
plot(pred_default, residuals(CCC05.gam),)
qqnorm(residuals(CCC05.gam))
plot(pred1, residuals(CCC05.gam1))
qqnorm(residuals(CCC05.gam1))
plot(pred2, residuals(CCC05.gam2))
qqnorm(residuals(CCC05.gam2))
plot(pred3, residuals(CCC05.gam3))
qqnorm(residuals(CCC05.gam3))
plot(pred4, residuals(CCC05.gam4))
qqnorm(residuals(CCC05.gam4))
```
From the summaries above, we can see that when spar had the value of 0.01, it provided the highest variance explained value, 54\%. This can also be seen from the plot above, the yellow line has the best fitting. From the residual plots, we can see that all these models had similar residual distributions. The Q-Q plots suggested there is a slight skewness in the residuals. 

```{r}
ECAN93 <- read.table("ECAN93.csv", header = TRUE, sep = ',', na.strings = "na")
ECAN93$Date <- as.Date(ECAN93$Date, "%d/%m/%Y")

ECAN93 <- transform(ECAN93, ndate = as.numeric(Date),
                   nyear  = as.numeric(format(Date, '%Y')),
                   nmonth = as.numeric(format(Date, '%m')),
                   nday = as.numeric(format(Date, '%j')))


# Start with the default model and 4 more with different spar parameters
ECAN93.gam <- gam(Value ~ s(nyear) + s(nmonth) +s(nday), data = ECAN93 )
pred_default <- predict(ECAN93.gam)
summary(ECAN93.gam)

ECAN93.gam1 <- gam(Value ~ s(nyear, sp=1.0) + s(nmonth, sp = 1.0) +
                    s(nday, sp = 1.0), data = ECAN93)
pred1 <- predict(CCC05.gam1)
summary(ECAN93.gam1)

ECAN93.gam2 <- gam(Value ~ s(nyear, sp=0.01) + s(nmonth, sp = 0.01) +
                    s(nday, sp = 0.01), data = ECAN93)
pred2 <- predict(ECAN93.gam2)
summary(ECAN93.gam2)

ECAN93.gam3 <- gam(Value ~ s(nyear, sp=0.25) + s(nmonth, sp = 0.25) +
                    s(nday, sp = 0.25), data = ECAN93)
pred3 <- predict(ECAN93.gam3)
summary(ECAN93.gam3)

ECAN93.gam4 <- gam(Value ~ s(nyear, sp=0.75) + s(nmonth, sp = 0.75) +
                    s(nday, sp = 0.75), data = ECAN93)
pred4 <- predict(ECAN93.gam4)
summary(ECAN93.gam4)

# ggplot(data = ECAN93, aes(x = Date, y = Value))
  # geom_point()
  # geom_line(aes(x=Date, y=pred_default), colour = "pink") +
  # geom_line(aes(x=Date, y=pred1), colour = "lightblue") +
  # geom_line(aes(x=Date, y=pred2), colour = "lightgreen") +
  # geom_line(aes(x=Date, y=pred3), colour = "purple") +
  # geom_line(aes(x=Date, y=pred4), colour = "black")
```


# QUESTION 2 Multiple Comparisons
```{r, warning=FALSE, message=FALSE}
# Read in dataset
library(xlsx)
herbicides <- read.xlsx("herbicides.xlsx", sheetIndex = 1)
```


(a) Carry out an analysis of variance on the data with Herbicide as the explanatory variable and Grass_percent" as the response.
```{r, warning=FALSE, message=FALSE}
library(multcomp)

Herbicides.aov <- aov(Grass_percent ~ Herbicide, herbicides)
summary(Herbicides.aov)
```

(b) Discuss the residuals
```{r}
# Residual distribution
par(mfrow = c(2, 2))
hist(Herbicides.aov$residuals)
qqnorm(Herbicides.aov$residuals)
plot(fitted(Herbicides.aov), residuals(Herbicides.aov))
abline(0,0)
```
From the histogram of residuals, we can see that it is pretty close to be normally distributed. The Q-Q plot also suggests that as there is no obvious skewness or tailed part. The points in the residual-fitted plot also are evenly distributed around 0 without any patterns. 

(c) Carry out an LSD type analysis comparing all possible pairs of treatments. Note which pairs have a significant difference.

```{r}
library(agricolae)
pairwise.t.test(herbicides$Grass_percent, herbicides$Herbicide, p.adj = "none")
mse <- sum(Herbicides.aov$residuals * Herbicides.aov$residuals)/Herbicides.aov$df.residual
LSD.test(herbicides$Grass_percent, herbicides$Herbicide, Herbicides.aov$df.residual, 
         mse, console = TRUE)
```
The LSD value obtained here is 7.14. From the result above, we can see that there is no significant difference between Aminopyralid and Aminopyralid + triclopyr, MCPA and Flumetsulam, Chlorsulfuron, MCPB+bentazone and Nil.

(d) Carry out pairwise comparisons using Bonferroni, Tukey and Dunnett adjustments and in each case show the pairs with significant differences.

```{r}
# Bonferroni adjustment
pairwise.t.test(herbicides$Grass_percent, herbicides$Herbicide, p.adj = "bonferroni",
                console = TRUE)

# Tukey adjustment
HerbicidesHSD <- TukeyHSD(aov(Grass_percent ~ Herbicide, herbicides))
plot(HerbicidesHSD)
summary(HerbicidesHSD)

# Dunnett adjustment
herbicides$HerbicideA <- as.factor(herbicides$Herbicide)
tdaov <- aov(Grass_percent~HerbicideA, data = herbicides)
test.dunnett = glht(tdaov, linfct = mcp(HerbicideA="Dunnett"))
confint(test.dunnett)
```


